Numerical simulations of the Cordilleran ice sheet through the last glacial cycle

### 2014-07-17 cordillera-cycle.tex
Rework sensitivity (results) part.

Despite large differences in the input atmospheric forcing (temperature offset
time-series; Fig.~\ref{fig:timeseries}), model output presents consistent
features that can be observed across the range of forcing data used. In all
simulations, modelled ice volumes remain relatively low during most of the
glacial cycle, except during two major glacial events which occur between 61.9
and 55.4\,ka during Marine Oxygen Isotope Stage (MIS)~4, and between 29.5 and
16.9\,ka during MIS~2 (Fig.~\ref{fig:timeseries}). An ice volume minimum is
consistently reached between 52.2 and 45.6\,ka during MIS~3. However, the
magnitude and precise timing of these three events depend significantly on the
choice of proxy record used for time-dependent atmospheric forcing
(Table~\ref{tab:extrema})\renote{
    Hopefully this new table will assess several of your comments at once.}.

Simulations forced by the Greenland ice core palaeo-temperature\renote{
    From here on, I prepend ``palaeo-temperature'' to ``record(s)'' in place of
    ``climate'' as you (Arjen) originally suggested. If we are to precise what
    type of ``records'' we use, let's be clear about it.}
records (GRIP, NGRIP) produce the highest variability in modelled ice volume
throughout the last glacial cycle. In contrast, simulations driven by oceanic
(ODP~1012, ODP~1020) and Antarctic (EPICA, Vostok) palaeo-temperature records
generally result in smaller\renote{
    Until we run out of other ajectives, I don't think that ``modest'' and
    ``generous'' are adequate terms for describing ice volumes.}
modelled ice volumes during MIS~4 and larger ice volumes during MIS~3.
More broadly speaking, they produce lower ice volume variability
throughout the simulation length. The NGRIP atmospheric forcing is the only one
that results in a larger ice volume during MIS~4 than during the LGM (MIS~2).

While simulations driven by GRIP and the two Antarctic palaeo-temperature
records attain a last ice volume maximum between 19.5 and 16.9\,ka, those
informed by NGRIP and the two oceanic palaeo-temperature records attain their
maximum ice volume thousands of years earlier. Moreover, the ODP~1012 run
yields a rapid deglaciation of the modelled area prior to 20\,ka. The ODP~1020
simulation predicts an early maximum in ice volume at 29.5\,ka, followed by
slower deglaciation than modelled using the other palaeo-temperature records.
Finally, model runs forced by Antarctic palaeo-temperature records results in a
rapid an uninterrupted deglaciation after the last glacial maximum, whereas the
simulation driven by the GRIP palaeo-temperature record also result in a rapid
deglaciation but in three steps, interrupted by two standstills and readvances
of the ice margin.

@@

Despite such differences in the timing of attained volume maxima and mimima,
all the model runs show relatively consistent patterns of glaciation
(Fig.~\ref{fig:snapshots}). As a result of scaling factors applied to the
different palaeo-temperature proxy records (Table~\ref{tab:records}), modelled
ice sheet geometries during the last glacial maximum (MIS~2;
Fig.~\ref{fig:snapshots}, lower panels) appear highly similar from one
simulation to the next. They include a ca. 1500\,km-long central divide
located about 3500\,m~a.s.l. along the spine of the Rocky
Mountains. The volumes of these modelled last glacial maximum ice sheets are
contained within a close range from 8.24 to 8.66\,m of sea-level equivalent
(s.l.e; Table~\ref{tab:extrema}).
A central ice cap persists in all simulations over the Skeena
Mountains during the MIS~3 ice volume minima (Fig.~\ref{fig:snapshots}, middle
panels). However, the size of the Skeena Mountains ice cap, and the dimensions
of the Cordilleran ice sheet during the MIS~4 glaciation, depend sensitively on
the choice of palaeo-temperature record used to drive the model.

\todo{Add general statements on the scale of glaciation.}

>>>>

Despite large differences in the input atmospheric forcing
(Fig.~\ref{fig:timeseries}, upper panel), model output presents consistent
features that can be observed across the range of forcing data used. In all
simulations, modelled ice volumes remain relatively low during most of the
glacial cycle, except during two major glacial events which occur between 61.9
and 55.4\,ka during MIS~4, and between 29.5 and 16.9\,ka during MIS~2
(Fig.~\ref{fig:timeseries}, lower panel). An ice volume minimum is
consistently reached between 52.2 and 45.6\,ka during MIS~3. However, the
magnitude and precise timing of these three events depend significantly on the
choice of proxy record used for time-dependent atmospheric forcing
(Table~\ref{tab:extrema})\renote{
    This new table contains modelled ages, volumes and extents at MIS~4--2}.

Simulations forced by the Greenland ice core palaeo-temperature\renote{
    From here on, I prepend ``palaeo-temperature'' to ``record(s)'' in place of
    ``climate'' as you (Arjen) originally suggested. If we are to precise what
    type of ``records'' we use, let's be clear about it.}
records (GRIP, NGRIP) produce the highest variability in modelled ice volume
throughout the last glacial cycle. In contrast, simulations driven by oceanic
(ODP~1012, ODP~1020) and Antarctic (EPICA, Vostok) palaeo-temperature records
generally result in smaller\renote{
    Until we run out of other ajectives, I don't think that ``modest'' and
    ``generous'' are necessary terms for describing ice volumes.}
modelled ice volumes during MIS~4 and larger ice volumes during MIS~3.
More broadly speaking, they produce lower ice volume variability
throughout the simulation length. The NGRIP atmospheric forcing is the only one
that results in a larger ice volume during MIS~4 than during the MIS~2.

While simulations driven by the GRIP and the two Antarctic palaeo-temperature
records attain a last ice volume maximum between 19.5 and 16.9\,ka, those
informed by the NGRIP and the two oceanic palaeo-temperature records attain their
maximum ice volume thousands of years earlier. Moreover, the ODP~1012 run
yields a rapid deglaciation of the modelled area prior to 20\,ka. The ODP~1020
simulation predicts an early maximum in ice volume at 29.5\,ka, followed by
slower deglaciation than modelled using the other palaeo-temperature records.
Finally, model runs forced by Antarctic palaeo-temperature records result in a
rapid and uninterrupted deglaciation after the last glacial maximum, whereas the
simulation driven by the GRIP palaeo-temperature record also results in a rapid
deglaciation but in three steps, separated by two ice volume standstills.

@@

Despite such differences in the timing of attained volume maxima and mimima
(Table~\ref{tab:extrema}), all the model runs show relatively consistent
patterns of glaciation (Fig.~\ref{fig:snapshots}, upper panels). During MIS~4,
all simulation produce an extensive ice sheet, covering an area of at least
half that attained during MIS~4. However, the corresponding maximum ice volumes
vary differ significantly between model runs, and cary between 3.88 and~8.69\,m
sea-level equivalent (m.~s.l.e; Table~\ref{tab:extrema}).

During the MIS~3 ice volume minima, a central ice cap persists in all
simulations over the Skeena Mountains (Fig.~\ref{fig:snapshots}, middle
panels). Although this ice cap is present in all simulations, its dimensions
depend sensitively on the choice of palaeo-temperature record used to drive the
model. Modelled ice volume minima vary between 0.89 and 2.44\,m~s.l.e.
(Table~\ref{tab:extrema}).

The modelled ice sheet geometries during the last glacial maximum (MIS~2;
Fig.~\ref{fig:snapshots}, lower panels) include a ca. 1500\,km-long central
divide located about 3500\,m~a.s.l. along the spine of the Rocky Mountains and
appear highly similar from one simulation to the next. Although the similatity
in modelled ice extent is a direct result from the choice of scaling factors
applied to the different palaeo-temperature proxy records
(Table~\ref{tab:records}), it is interesting to note that modelled maximum ice
volumes are also contained within a close range of 8.24 to 8.66\,m~s.l.e.
(Table~\ref{tab:extrema}).
